---
title: "Web Scraping v2"
author: "STAT 345 - Statistical Computing"
output: ioslides_presentation
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(out.width = '80%', collapse = TRUE, warning=FALSE, message=FALSE)
library(tidyverse)
```

## Agenda

- HTML and CSS 
- `rvest` for scraping
- String basics with `stringr`

Readings: Irizarry 23-24

## Murders

The original data for the `murders` dataset comes from [https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state](https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state)

Right click, then View Page Source to see the HTML

## HTML

It's worth learning some HTML and CSS coding!

- Personal websites to showcase your work
- Better applications (dashboards and Shiny apps)

## HTML

```{r, eval = FALSE}
</th></tr></tbody></table>
<table class="wikitable sortable">
<tbody><tr>
<th data-sort-type="text">State
</th>
<th data-sort-type="number">Population<br /> <small>(total inhabitants)</small> <br /> <small>(2015)</small> <sup id="cite_ref-2" class="reference"><a href="#cite_note-2">&#91;2&#93;</a></sup>
</th>
<th data-sort-type="number">Murders and<br />Nonnegligent<br />Manslaughter<br /> <small>(total deaths)</small> <br /> <small>(2015)</small> <sup id="cite_ref-Crime_1-1" class="reference"><a href="#cite_note-Crime-1">&#91;1&#93;</a></sup>
</th>
<th data-sort-type="number">Murders<br /> <small>(total deaths)</small> <br /> <small>(2015)</small> <sup id="cite_ref-Murders_3-0" class="reference"><a href="#cite_note-Murders-3">&#91;3&#93;</a></sup>
</th>
<th data-sort-type="number">Gun Murders<br /> <small>(total deaths)</small> <br /> <small>(2015)</small> <sup id="cite_ref-Murders_3-1" class="reference"><a href="#cite_note-Murders-3">&#91;3&#93;</a></sup>
</th>
<th data-sort-type="number">Gun<br />Ownership<br /> <small>(%)</small> <br /> <small>(2013)</small> <sup id="cite_ref-4" class="reference"><a href="#cite_note-4">&#91;4&#93;</a></sup>
</th>
<th data-sort-type="number">Murder and<br />Nonnegligent<br />Manslaughter<br />Rate<br /> <small>(per 100,000)</small> <br /> <small>(2015)</small>
</th>
<th data-sort-type="number">Murder Rate<br /> <small>(per 100,000)</small> <br /> <small>(2015)</small>
</th>
<th data-sort-type="number">Gun<br />Murder Rate<br /> <small>(per 100,000)</small> <br /> <small>(2015)</small>
</th></tr>
<tr>
<td><a href="/wiki/Alabama" title="Alabama">Alabama</a>
</td>
<td style="background: #ffddcc">4,853,875
</td>
<td style="background: #ffddcc">348
</td>
<td>—<sup id="cite_ref-limited-data_5-0" class="reference"><a href="#cite_note-limited-data-5">&#91;a&#93;</a></sup>
</td>
<td>—<sup id="cite_ref-limited-data_5-1" class="reference"><a href="#cite_note-limited-data-5">&#91;a&#93;</a></sup>
</td>
<td style="background: #ffbbcc">48.9
</td>
<td style="background: #ffbbcc">7.2
</td>
<td>— <sup id="cite_ref-limited-data_5-2" class="reference"><a href="#cite_note-limited-data-5">&#91;a&#93;</a></sup>
</td>
<td>— <sup id="cite_ref-limited-data_5-3" class="reference"><a href="#cite_note-limited-data-5">&#91;a&#93;</a></sup>
</td></tr>
<tr>  
```

## rvest

- `rvest` is a web scraping package in the `tidyverse`. 
- The first step in using this package is to import the web page into R:

```{r, message=FALSE, warning=FALSE}
library(rvest)
url <- "https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state"
h <- read_html(url)
```

## rvest

```{r}
class(h)
h
```

## Nodes

The different parts of an HTML document, often defined with a message in between `<` and `>` are referred to as __nodes__. 

- The `rvest` package includes functions to extract nodes of an HTML document: 
  - `html_nodes` extracts all nodes of different types, and 
  - `html_node` extracts the first one.
  
## Nodes

```{r}
tab <- h %>% html_nodes("table")
class(tab)
tab
```

We actually just want the object in `tab` that defines the wikitable. 

## Nodes to tables

```{r}
tab <- h %>% html_nodes("table") %>% .[2]
class(tab)
```

## Nodes to tables

`rvest` includes a function just for converting HTML tables into data frames:

```{r}
tab <- tab %>% html_table %>% .[[1]] 
class(tab)
```

## Nodes to tables

This is pretty close! There's still some work to do to clean this up, though!

```{r}
tab <- tab %>% setNames(c("state", "population", "murder_manslaughter_total", "murder_total", "gun_murder_total", "ownership", "murder_manslaughter_rate",  "murder_rate", "gun_rate"))
head(tab)
```

## CSS Selectors

- HTML is generally unattractive, and most websites use CSS
- CSS controls the way elements on the webpage look
  - titles, headings, lists, tables, links, etc.
  - The patterns used here are called _selectors_
- If we know the selector, we can use `html_nodes`
  - This is tricky, and getting trickier
  
## SelectorGadget

This is a browser add-on (Chrome extension) that is useful for finding the selector you need to extract data.

For example, imagine we want to find the actors listed on an IMDB movie page, e.g. The Lego Movie.

- Navigate to the page and find the actors list
- Turn on the Selector gadget
- Click on the element you want to select, it will turn green (and guesses the css selector)
  - All matches are highlighted yellow
- Click on elements you don't want -- they'll turn red
- Copy the css selector returned by SelectorGadget

## Extracting the values

```{r}
lego_url <- "https://www.imdb.com/title/tt1490017/fullcredits/?ref_=tt_cl_sm"
html <- read_html(lego_url)
cast <- html_nodes(html, ".primary_photo+ td a")
length(cast)
cast[1:2]
```

## Common node functions

- `html_tag()`: extract tag names
- `html_text()`: extract text
- `html_attr()`: extract single attribute 
- `html_attrs()`: extract all attributes

## Extract the text

```{r}
html_text(cast, trim = TRUE)
```

## html attributes

Let’s say we’re also interested in extracting the links to the actors’ pages. We can access html attributes of the selected nodes using `html_attrs()`.

```{r}
cast_attrs <- html_attrs(cast)
length(cast_attrs)
cast_attrs[1:2]
```

## html attributes

- One attribute called `href` which contains relative url to the actor’s page. 
- Extract it using `html_attr()`, indicating the name of the attribute of interest. 
- Relative urls can be turned to absolute urls using `url_absolute()`

## html attributes

```{r}
cast_abs_urls <- html_attr(cast, "href") %>% 
  url_absolute(lego_url)
cast_abs_urls[1:2]
```

## Example - guacamole

Extract the recipe name, total preparation time, and list of ingredients from [this](http://www.foodnetwork.com/recipes/alton-brown/guacamole-recipe-1940609) guacamole recipe.

```{r}
h           <- read_html("http://www.foodnetwork.com/recipes/alton-brown/guacamole-recipe-1940609")
recipe      <- h %>% html_node(".o-AssetTitle__a-HeadlineText") %>% html_text()
prep_time   <- h %>% html_node(".m-RecipeInfo__a-Description--Total") %>% html_text()
ingredients <- h %>% html_nodes(".o-Ingredients__a-Ingredient") %>% html_text()
```

## Make the guacamole

```{r}
guacamole <- list(recipe, prep_time, ingredients)
guacamole
```

## Get more recipes, faster

```{r}
get_recipe <- function(url){
  h           <- read_html(url)
  recipe      <- h %>% html_node(".o-AssetTitle__a-HeadlineText") %>% html_text()
  prep_time   <- h %>% html_node(".m-RecipeInfo__a-Description--Total") %>% html_text()
  ingredients <- h %>% html_nodes(".o-Ingredients__a-Ingredient") %>% html_text()
  return(list(recipe = recipe, prep_time = prep_time, ingredients = ingredients))
}
```

## Get more recipes

```{r}
get_recipe("http://www.foodnetwork.com/recipes/food-network-kitchen/pancakes-recipe-1913844")
```


## String basics

```{r}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

## String basics

```{r}
double_quote <- "\"" # or '"'
single_quote <- '\'' # or "'"
```

## String basics

```{r}
x <- c("\"", "\\")
x
#> [1] "\"" "\\"
writeLines(x)
```

Other characters: try `?'"'`
  
## String operations

String processing tasks can be divided into:

- detecting, 
- locating, 
- extracting, or 
- replacing

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
| `str_detect` | Detect | Is the pattern in the string? | `grepl` |
| `str_which` | Detect | Returns the index of entries that contain the pattern. |  `grep` |
| `str_subset` | Detect | Returns the subset of strings that contain the pattern. | `grep` with `value = TRUE` |
| `str_locate` | Locate | Returns positions of first occurrence of pattern in a string.| `regexpr` |
|  `str_locate_all` | Locate | Returns position of all occurrences of pattern in a string. |`gregexpr` |
| `str_view` | Locate | Show the first part of the string that matches pattern. | |

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
 | `str_view_all` | Locate | Show me all the parts of the string that match the pattern. | 
| `str_extract` | Extract | Extract the first part of the string that matches the pattern. | |
`str_extract_all` | Extract | Extract all parts of the string that match the pattern. |  |
 | `str_match` | Extract | Extract first part of the string that matches the groups and the patterns defined by the groups.| |
| `str_match_all` | Extract | Extract all parts of the string that matches the groups and the patterns defined by the groups. | |
| `str_sub` | Extract | Extract a substring. | `substring` |

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
| `str_split` | Extract | Split a string into a list with parts separated by pattern. | `strsplit`|
| `str_split_fixed` | Extract | Split a string into a matrix with parts separated by pattern. | `strsplit` with `fixed = TRUE`|
| `str_count` | Describe | Count number of times a pattern appears in a string.  |   |
| `str_length`| Describe | Number of character in string.  | `nchar` |
| `str_replace` | Replace | Replace first part of a string matching a pattern with another.  | | 

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
| `str_replace_all`| Replace | Replace all parts of a string matching a pattern with another.  | `gsub` | 
 | `str_to_upper` | Replace | Change all characters to upper case. | `toupper` | 
| `str_to_lower` | Replace | Change all characters to lower case. | `tolower` |
| `str_to_title` | Replace | Change first character to upper and rest to lower. | | 
 | `str_replace_na`| Replace | Replace all `NA`s to a new value. | |
| `str_trim` | Replace | Remove white space from start and end of string. | |

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
| `str_c`| Manipulate | Join multiple strings.  | `paste0` |
 | `str_conv` | Manipulate | Change the encoding of the string.| |
 | `str_sort` | Manipulate | Sort the vector in alphabetical order.| `sort` |
 | `str_order`| Manipulate | Index needed to order the vector in alphabetical order. | `order` |
| `str_trunc` | Manipulate | Truncate a string to a fixed size. | |
| `str_pad`| Manipulate | Add white space to string to make it a fixed size.  | |

## `stringr` functions

| stringr     | Task     | Description                              | R-base |
|--------------- |----------|--------------------------------------|-------------|
| `str_dup`| Manipulate | Repeat a string.  | `rep` then `paste` |
 | `str_wrap` | Manipulate | Wrap things into formatted paragraphs. | |
| `str_interp` | Manipulate | String interpolation. |  `sprintf` | 

## String length

```{r}
str_length(c("a", "Survey of Modern Statistical Software", 9))
```

## String concatenation

```{r}
str_c("x", "y")
str_c("x", "y", "z")
str_c("x", "y", sep = ", ")
```

## String concatenation

```{r}
x <- c("abc", NA)
str_c("|-", x, "-|")
str_c("|-", str_replace_na(x), "-|")
```

## String subsetting

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1)
```

## Raw murders

```{r,message=FALSE, warning=FALSE}
library(rvest)
url <- paste0("https://en.wikipedia.org/w/index.php?title=",
              "Gun_violence_in_the_United_States_by_state",
              "&direction=prev&oldid=810166167")
murders_raw <- read_html(url) %>%
  html_node("table") %>%
  html_table() %>%
  setNames(c("state", "population", "total", "murder_rate"))
```

## Raw murders

Converting to numeric from text (even nicely formatted text) is trickier.

```{r}
murders_raw$population[1:3]
```

The usual coercion does not work here:

```{r}
as.numeric(murders_raw$population[1:3])
```

## Detecting commas

We can use the `str_detect` function to see that two of the three columns have commas in the entries:

```{r}
commas <- function(x) any(str_detect(x, ","))
murders_raw %>% summarize_all(commas)
```

## Removing commas

We can then use the `str_replace_all` function to remove them:

```{r}
test_1 <- str_replace_all(murders_raw$population, ",", "")
test_1 <- as.numeric(test_1)
test_1
```

## readr version

```{r}
test_2 <- parse_number(murders_raw$population)
identical(test_1, test_2)
```
